Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-09-17 doc/paper-lcr.tex
edit draft

If the mass sheet is not physically associated with the lens galaxy, it can lie at any redshift. The effects of such a mass-sheet can be projected back onto the lens plane in the form of an effective convergence. We define this effective convergence in two limits.
\begin{itemize}
\item No significant perturbations along the line of sight, and \\
\item A single dominant perturber, relative to which all other perturbations are small.
\end{itemize}
We will refer to these cases as the magnification case (subscript $\mu$ in formulae), and the strong lensing case (subscript SL in formulae), respectively. We will ignore the additional case, of a compound lens, with several significant perturbers.

@@

In the stong lensing case, the presence of a significant perturber alters the
geometry of the system, and provides a primary lens plane. It is the primary
lens plane upon which we need to know $\kappa_{\mathrm{effective}}$, since
this is where the mass-sheet degeneracy is defined. Following Equations 20--32
of \citet{keeton2003}\flag{keeton2003}, it can be shown that the relevant
effective convergence is given by 
\be \label{eq:kappasumSL}
\kappa_{\mathrm{effective,SL}}=
{  
(1-\beta_{12})(\kappa-\beta_{12}(\kappa^2-\gamma^2))
\over
(1-\beta_{12}\kappa)^2-(\beta_{12}\gamma)^2
}
\ee
where $\kappa$ and $\gamma$ are, respectively, the convergence and shear
caused by the perturber, in their own plane and $\beta_{12}$ is a geometrical
scaling factor defined by
\be\label{eq:beta}
\beta_{12}={
D_{12} D_{\mathrm{os}}
\over
D_{\mathrm{o}1} D_{1\mathrm{s}}
}
\ee
where the D's are again angular diameter distances: the 1 (2) in the subscript
refers to the lower (higher) redshift object of the lens and perturber. For
the case that the mass sheet is in the lens plane, $\beta_{12} = 0$ and
$\kappa_{\mathrm{eff}}=\kappa$; if the mass-sheet is in the plane of the
observer, or source, $\beta_{12} = 1$ and $\kappa_{\mathrm{eff}}=0$; for a
mass-sheet at any other redshift $0<\beta_{12}<1$. In the case of several mass
sheets the effective convergences sum. In the limit that both $\kappa$ and
$\gamma$ are small, Equation \ref{eq:kappasumSL} reduces to 
\be \label{eq:kappasumSLweak}
\kappa_{\mathrm{effective,SL}}=(1-\beta_{12})\kappa;
\ee
it is this $\kappa_{\mathrm{effective,SL}}$ that we use as $\kappax$ for the
strong lensing time delay analysis in \ref{sec:Dt}.

@@

\subsection{Voids}

Whilst halos contribute a positive convergence to $\kappax$, voids contribute
a negative convergence. When a raybundle passes through a  region of space
that is less dense than $\rho_{\rm cr}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos, is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full raytraced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. In principle the absence of a halo could be taken to infer a
void, but the under-density of such a void is hard to infer. Instead, we shift
our $\kappax$ values so that $\langle\kappax\rangle=0$. We refer to this
corrective term as $\kappav$. The void correction introduces a source of
error, since the inferred value of $\kappav$ will depend on the properties of
the sightlines reconstructed, but this error shrinks with the number of
sightlines reconstructed. To avoid potential biases we calculate $\kappav$
from randomly selected sightlines. The total $\kappax$ along the line of sight
is given by 
\be \label{eq:totalkappax}
\kappax = \left(\sum_{\rm halos} \kappa_{\rm halo}\right)-\kappav.
\ee

\flag{Need to draw a convergence from the void distribution, such that 1) the
intrinsic distribution of kappas is returned in the case where we have no
data, and 2) the mean density parameter over all lines of sight is $\Omega_m$.
The mass budget has to be balanced!}

@@

Given the halo mass catalogue from the Millenium Simulation, and convergence maps produced by
ray-tracing through the MS's n-body particles, 
it is possible to test the validity of the prescription outlined in Section \ref{sec:model}. We do this
using $\kapparay$ as a proxy for $\kappatrue$. The reconstruction prescription is tested given perfect
knowledge of the halos' virial masses and redshifts; this provides us with a measure of the intrinsic
errors induced by treating matter as truncated NFW halos. To quantify this 10000 lines of sight are chosen
from random positions in the catalogue, and for each line of sight we reconstruct the
convergences from every halo within 5 arcminutes and compare the reconstructed
kappas with those from the ray tracing results. The standard deviation on $\kapparay$-$\kapparec$ 
shown in Figure \ref{fig:ScattervsTruncation}. It is clear from Figure \ref{fig:ScattervsTruncation} 
that as long as the truncation is greater than a few virial radii, the choice of truncation
radius does not broadly affect the scatter on $\kapparay-\kapparec$; from now on we choose 
to truncate the halos at $5 \Rhalo$. Since $\sigma\left(\kapparay-\kapparec\right)$ does not go to
zero, it is clear that out halo based prescription does not perfectly replicate the n-body results, we refer to this
 as the intrinsic error of the prescription. The intrinsic error has a standard deviation of $\sigi \sim 0.018$, although
this figure is inflated by a few outliers in the tails of $\pr\left(\kapparay-\kapparec\right)$. Whilst we do not
focus on shear in this work, such outliers would likely be rejected by their reconstructed shear not matching the
shear of the lens model (see \citet{wong} for the practicalities of this), this makes our estimate of the intrinsic error pessimistic.

\begin{figure}
\includegraphics[width=\columnwidth]{truncation_scatter.eps}
\caption[magcut]{}
\label{fig:ScattervsTruncation}
\end{figure}

\tom{this paragraph is phrased badly}
If $\kapparec$ is to be used in a real survey it is important that it is not a biased estimator of the true $\kappax$. 
Ideally one wants $\langle \kapparec-\kappatrue \rangle$ to equal 0 and $\kapparec-\kappatrue$ to be independant of
$\kapparec$., since the void correction ensures $\langle \kapparec \rangle = 0$ any bias in $\langle \kapparec-\kapparay \rangle$ comes
from the variations in $\kapparay$. With 8.2$\times 10^8$ pixels, uniformly spaced over 196 deg$^2$ of simulated sky,
 $\langle \kapparay \rangle$ = -0.00014. However, it is improbable that such a large number of lines of sight
 could be reconstructed in reality and the mean of $\kappatrue$ for a small subset of sightlines will not necessarily
equal zero. We find this error to be
\be\label{eq:sigmav}
\sigma(\kappav) \simeq 0.036 / N^{\frac{1}{2}},
\ee
 where $N$ is the number of lines of sight, although $\pr(\kappav)$ has positive skew for small $N$.
 Care must be taken that the callibration lines of sight are not biased in $\kappax$ or the final $\pr(\kappav)$ will also be biased. For the rest of this work we use 1000  randomly chosen calibration lines of sight to calculate $\kappav$, introducing an error on $\kappax$ of $0.001$. 1000 callibration lines of sight is unfeasable currently, but should be reasonable in the LSST era. Since the true $\langle \kapparay \rangle$ = -0.00014 is two orders of magnitude smaller than intrinsic error of the halo prescription, $\kapparec$ can be seen as unbiased given a large enough sample of lines of sight. However since $\kappax$ reconstruction based on perfect knowledge of the halo mass and position is unrealistic we must re-visit the question of bias in Section \ref{sec:obsMstar+z}, using more realistic observables.

\comment{
 and Figure \ref{fig:bias} shows that there is no statistically significant evidence that $\kapparec$ is a biased estimator, regardless of the individual value of $\kapparec$.

\begin{figure}
\includegraphics[width=\columnwidth]{isitbiased.eps}
\caption[Biased?]{}
\label{fig:isitbiased}
\end{figure}
}
\begin{figure}
\includegraphics[width=\columnwidth]{mag_scatter.eps}
\caption[magcut]{}
\label{fig:magcut}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{radius_scatter.eps}
\caption[radius cut]{}
\label{fig:radcut}
\end{figure}

The error from the void correction, and the intrinsic error from the halo based prescription are unavoidable errors, even with perfect knowledge of halo mass and position. Before investigating the errors caused by imperfect knowledge of halo mass and redshift, we investigate the errors induced by a magnitude limited reconstruction and a field of view limited reconstruction. The halos in our catalogue are given magnitudes by the semi-analytic model of \citet{SAM} and by applying magnitude cuts to our catalogue, we can investigate the scatter caused by unobserved halos. The majority of the convergence comes from halos with an $i$ magnitude between 18 and 24. Figure \ref{fig:magcut} shows that the width of $\kapparec-\kapparay$ decreases quickly between $i=18$ and $i=24$. Objects brighter than $i=18$ are either too rare or too close to the observer to make a significant contribution to the convergence. Objects fainter than $i=24$ are too small to be important unless they are extremely close to the line of sight where it is likely that neglecting stellar mass and using a spherical NFW prescription are too naive to adequately reconstruct $\kappax$ - it is likely that these ultra-faint halos (or substructures) are the major source of the scatter in $\kapparec-\kapparay$, but a deeper survey will not be sufficient to decrease the size of this scatter. Since a large number of callibration lines of sight are needed to shrink the width of $\pr(\kappav)$, it may prove unrealistic to reconstruct the calibration lines of sight out to 5 arcminutes. Figure \ref{fig:radcut} shows the width of $\kapparec-\kapparay$ as a function of reconstruction radius; in almost all cases there is no significant contribution to $\kappa$ from halos that are centred more than 1 arcminute away, although large groups or clusters can sometimes still make a contribution as far out as four arcminutes. For the rest of this work we will continue to model all of the halos out to 5 arminutes with an $i$ magnitude greater than 26, although a reconstruction with fields going out to two arcminutes and down to $i=24$ would do almost as well.

@@

Real attempts to reconstruct the convergence along a line of sight will not have direct access to the masses of dark matter halo, 
and will likely not have access to their spectroscopic redshift either. Instead these properties must be inferred from astrophysical
observables. In principle the only observables are astrometric positions, photometric colours and possibly spectra. In this section we
quantify the errors induced by inferring the halo mass and redshift from the observables. Much work has already focused on using
photometric colours to infer stellar mass \citep[\eg][]{xxx} and redshifts \citep[\eg][]{BPZ}, in line with these we shall use
stellar mass and photometric redshifts (with appropriate errors), as the observables that must be converted into an estimate
of $\kappax$. We will investigate three main sources of error: inferring halo mass given observed stellar mass; halos at the wrong redshift
due to photometric redshift error, and; the combination of both redshift and stellar mass errors.

We generate a stellar mass for each halo according to the stellar mass--halo mass relation of \citet{behroozi2010}. From this stellar mass we simulate 
an observed stellar mass by drawing a sample from $\pr(\log(M_{* \mathrm {obs}})|\log(M_{* \mathrm {true}}))$ which is given by the product of the galaxy stellar mass function (GSMF) of \citet{Fontana2006} and a Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where a spectroscopic redshift exists stellar masses can be estimated with typical uncertainties of 0.15 dex \citep{xxx}, however with photometric redshifts stellar mass uncertainties are typically three times as large \citep{xxx}; we use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and $\sigma_{M_*}=0.45$ otherwise. The \citet{Fontana2006} GSMF is given by a Schechter function with redshift evolving parameters
\be
\label{eq:GSMF}
{\dee N \over \dee M_*} \propto \left(10^{(M_*-M_0(z))}\right)^{(1+\alpha(z))} \exp \left(-10^{(M_*-M_0(z))}\right)
\ee 
where $M_0(z)=11.16+0.17z-0.07z^2$ and $\alpha(z)=-1.18-0.082z$. For photometric redshift errors we draw a redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take as a Gaussian of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$ is the halo's true redshift in \MS catalogue

\begin{figure}
\includegraphics[width=\columnwidth]{widthvkappa.eps}
\caption[radius cut]{}
\label{fig:width}
\end{figure}

Our results are summarized in Figure \ref{fig:width}. We reconstruct 10000 lines of sight and each of the reconstructed $\pr(\kappa)$ is summarized by their mean $\kapparec$, and their width containing $68\%$ of the probability, $\sigt$. Reconstructing the lines of sight with perfect knowledge of the redshift but an uncertain stellar mass, we find that $\sigt$ grows with $\kapparec$; this is not surprising since the low $\kapparec$ lines of sight are relatively empty and so there are few opportunities for uncertainties in the halo masses to propogate into $\kapparec$ uncertainties. Our results are well fit by a straight line given by 
\be
\sigt=0.012+0.26 \langle\kapparec\rangle;
\label{eq:Mstarfit}
\ee
this is the reconstruction error that could be expected from a complete spectroscopic survey of the field. A complete spetroscopic survey is unrealistic, but Equation \ref{eq:Mstarfit} represents a lower limit on the reconstruction error.
We find that with a perfect knowledge of the halo mass, but photometric redshifts the reconstructed $\pr(\kappa)$ has a mean $\sigt$ of 0.003. For the lines of sight with $\kapparec {\rm s}$ that are dominated by a single significant perturber the photometric redshift error can induce a larger uncertainty on $\kapparec$, but only $2.9 \%$ of sightlines have $\sigt > 0.01$; typical photometric redshift errors do not propogate into a significant uncertainties on $\kapparec$ if the halo mass is known. This is seen in Figure \ref{fig:width}, $\sigt$ is typically much larger for the perfect redshift reconstructions than the perfect mass reconstructions. The results are well fit by
\be
\sigt=0.0025+0.041\langle\kapparec\rangle
\label{eq:photozfit}
\ee
Inferring the stellar mass of a galaxy from its magnitude and colours requires an estimate of how far away the galaxy is; without a spectroscopic redshift the the infered stellar mass is less precise. In principle the photometric redshift is correlated with the infered stellar mass however we do not model this since the redshift error produces a small effect on the recovered $\pr(\kappa)$ when compared to the effect of uncertain stellar masses.  For the reconstructions with imperfect redshift and stellar mass the resulting uncertainty is significantly larger than the quadrature sum of the $\sigt {\rm s}$ in Equations \ref{eq:Mstarfit} and \ref{eq:photozfit}, this is because the uncertain redshift propogates into a broadened $\pr(M_{*})$. The  
ns than the perfect mass reconstructions. For a complete photometric survey down to $i=26$ our reconstructions are fit by the line
\be
\sigt=0.026+0.037\langle\kapparec\rangle
\label{eq:bothfit}
\ee

>>>>

\comments{
If the mass sheet is not physically associated with the lens galaxy, it can lie at any redshift. The effects of such a mass-sheet can be projected back onto the lens plane in the form of an effective convergence. We define this effective convergence in two limits.
\begin{itemize}
\item No significant perturbations along the line of sight, and \\
\item A single dominant perturber, relative to which all other perturbations are small.
\end{itemize}
We will refer to these cases as the magnification case (subscript $\mu$ in formulae), and the strong lensing case (subscript SL in formulae), respectively. We will ignore the additional case, of a compound lens, with several significant perturbers.

@@

In the stong lensing case, the presence of a significant perturber alters the
geometry of the system, and provides a primary lens plane. It is the primary
lens plane upon which we need to know $\kappa_{\mathrm{effective}}$, since
this is where the mass-sheet degeneracy is defined. Following Equations 20--32
of \citet{keeton2003}\flag{keeton2003}, it can be shown that the relevant
effective convergence is given by 
\be \label{eq:kappasumSL}
\kappa_{\mathrm{effective,SL}}=
{  
(1-\beta_{12})(\kappa-\beta_{12}(\kappa^2-\gamma^2))
\over
(1-\beta_{12}\kappa)^2-(\beta_{12}\gamma)^2
}
\ee
where $\kappa$ and $\gamma$ are, respectively, the convergence and shear
caused by the perturber, in their own plane and $\beta_{12}$ is a geometrical
scaling factor defined by
\be\label{eq:beta}
\beta_{12}={
D_{12} D_{\mathrm{os}}
\over
D_{\mathrm{o}1} D_{1\mathrm{s}}
}
\ee
where the D's are again angular diameter distances: the 1 (2) in the subscript
refers to the lower (higher) redshift object of the lens and perturber. For
the case that the mass sheet is in the lens plane, $\beta_{12} = 0$ and
$\kappa_{\mathrm{eff}}=\kappa$; if the mass-sheet is in the plane of the
observer, or source, $\beta_{12} = 1$ and $\kappa_{\mathrm{eff}}=0$; for a
mass-sheet at any other redshift $0<\beta_{12}<1$. In the case of several mass
sheets the effective convergences sum. In the limit that both $\kappa$ and
$\gamma$ are small, Equation \ref{eq:kappasumSL} reduces to 
\be \label{eq:kappasumSLweak}
\kappa_{\mathrm{effective,SL}}=(1-\beta_{12})\kappa;
\ee
it is this $\kappa_{\mathrm{effective,SL}}$ that we use as $\kappax$ for the
strong lensing time delay analysis in \ref{sec:Dt}.
}

@@

\subsection{Calibration: Recovering the global $\pr(\kappax)$ using a void correction}
\label{subsec:voids}

Whilst halos contribute a positive convergence to $\kappax$, voids contribute
a negative convergence. When a raybundle passes through a  region of space
that is less dense than $\rho_{\rm cr}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos, is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full raytraced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. In principle the absence of a halo could be taken to infer a
void, but the under-density of such a void is hard to infer. Neglecting voids
would lead to a heavily biased estimate of $\kappax$. We account for the voids
statistically; we reconstruct
the $\kappax$ contribution of halos for an ensemble of calibration sightlines
and form the joint distribution P($\kapparay,\sumkappah$). The calibration
can then be accounted for by marginalizing over all lines of sight with a similar
 $\sumkappah$.  We will refer to this calibration proceedure as the "void correction",
although it also accounts for all of the physics that our simple spherical halo prescription
neglects in a statistical manner. For each line of sight we create a $\pr(\kappax)$ from
\be
\pr(\kappax)=\int_{-\infty}^{+\infty}\pr(\kapparay',\tilde\kappa'_{\rm halos})W(\tilde\kappa_{\rm halos}-\tilde\kappa'_{\rm halos})\dee \tilde\kappa'_{\rm halos}
\ee
where $\tilde{\kappa}_{\rm halos}$ is the median of the reconstructed $\sumkappah$ for each line of sight.
Primed qunatities refer to all of the calibration sightlines and unprimed refers to the specific line of sight.
$W(\tilde\kappa_{\rm halos}-\tilde\kappa'_{\rm halos})$ is a weight function given by
\be
W(\tilde\kappa_{\rm halos}-\tilde\kappa'_{\rm halos})=\exp\left(-\frac{1}{2}\left(\frac{\left(\tilde\kappa_{\rm halos}-\tilde\kappa'_{\rm halos}\right)^2}{\sigma^2}\right)\right)
\ee
We take $\sigma$ as 0.01- this is an overestimate of the uncertainty on $\tilde\kappa_{\rm halos}$, but ensures that the full distibution of void corrections is sampled, despite the finite number of calibration sightlines.

\comments{
For each reconstruction of a line of sight we can then draw a $\kappav$ from $\pr(\kappav)$ to correct for the
unseen voids. We refer to this corrective shift as $\kappav$. The void correction introduces a source of
error, since the value of $\kappav$ comes from the ensemble of calibration lines of sight rather than the individual
line of sight which is being reconstructed; this is unavoidable without modeling the $\kappax$ 
contribution of voids along a line of sight. To avoid potential biases we calculate $\pr(\kappav)$
from randomly selected calibration sightlines. The total $\kappax$ along the line of sight
is given by 
\be \label{eq:totalkappaxunweighted}
\pr(\kappax) = \left(\sum_{\rm halos} \pr(\kappa_{\rm halo})\right)-\pr(\kappav).
\ee

Once the void correction has been applied,the reconstruction should produce a 
reasonable estimate of $\pr(\kappax)$ for any single line of sight. However, from
the ray tracing results we know the global $\kappax$ distribution ${\rm U}(\kappax)$. We can use the global
 ${\rm U}(\kappax)$ as a prior by noting that the normalized sum of $\pr(\kapparec)$ from multiple lines
of sight should equal ${\rm U}(\kappax)$, \ie
\be \label{eq:prior}
{\rm U}(\kappax) = \lim_{N \to \infty} \sum_{i}^{N} \pr(\kappax)_{i} / N
\ee
In order to make Equation \ref{eq:prior} hold, we use Equation \ref{eq:totalkappaxunweighted} to calculate $\pr(\kappax)^{\rm unweighted}$ for the calibration lines of sight and
re-weight the reconstructions by:
\be \label{eq:weighting}
W(\kappa)=\left(\sum_{i}^{N} \pr(\kappax)_{i}^{\rm unweighted} /N \right) / \left({\rm U}(\kappax)\right)
\ee
such that the final reconstruction is given by 
\be \label{eq:totalkappax}
\pr(\kappax) = \left(\left(\sum_{\rm halos} \pr(\kappa_{\rm halo})\right)-\pr(\kappav)\right)(W(\kappa)).
\ee

}

\comments{
Whilst halos contribute a positive convergence to $\kappax$, voids contribute
a negative convergence. When a raybundle passes through a  region of space
that is less dense than $\rho_{\rm cr}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos, is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full raytraced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. In principle the absence of a halo could be taken to infer a
void, but the under-density of such a void is hard to infer.
\tom{ Instead, we reconstruct 10000 lines of sight to give $\pr(\kapparec|kapparay)$
and invert the PDF using Bayes' theorem to give $\pr(\kapparay|\kapparec)$. For each lightcone we
can then get $\pr(\kapparay)$ by marginalizing over the reconstruction PDF, $\pr(\kapparec)$.
This recipe also ensures that we correctly recover the ensemble $\kapparay$ distribution, even in the case
of no input data.

 Instead, we shift
our $\kappax$ values so that $\langle\kappax\rangle=0$. We refer to this
corrective term as $\kappav$. The void correction introduces a source of
error, since the inferred value of $\kappav$ will depend on the properties of
the sightlines reconstructed, but this error shrinks with the number of
sightlines reconstructed. To avoid potential biases we calculate $\kappav$
from randomly selected sightlines. The total $\kappax$ along the line of sight
is given by 
\be \label{eq:totalkappax}
\kappax = \left(\sum_{\rm halos} \kappa_{\rm halo}\right)-\kappav.
\ee
}}

@@

Given the halo mass catalogue from the Millenium Simulation, and convergence maps produced by
ray-tracing through the MS's n-body particles, 
it is possible to test the validity of the prescription outlined in Section \ref{sec:model}. We test the model
 by comparing $\sumkappah$ against $\kapparay$. If the reconstruction process is to be
 succesfull the two quantities should correlate.

The reconstruction prescription is tested given perfect
knowledge of the halos' virial masses and redshifts; this provides us with a measure of the intrinsic
errors induced by treating matter as truncated NFW halos. In this Section we do not apply the void correction of
Section \label{subsec:voids}, instead we shift $\sumkappah$ for all our sightlines by a fixed negative offset such that the ensemble mean
is zero, i.e.
\be \label{eq:totalkappax}
\kapparec = \left(\sum_{\rm halos} \kappa_{\rm halo}\right)-\kappav,
\ee
where $\kappav$ is a positive constant. This us allows us to test the halo reconstruction directly, without the uncertainty of the void correction
obscuring any inadequacies of our halo based prescription.

We choose 100000 lines of sight from random positions in the catalogue, and for each line of sight we reconstruct the
convergences from every halo within 5 arcminutes and compare the reconstructed
$\kapparec$ with those from the ray tracing results. The standard deviation on $\kapparay$-$\kapparec$ 
shown in Figure \ref{fig:ScattervsTruncation} as a function of the radius at which the halo is truncated.
 It is clear from Figure \ref{fig:ScattervsTruncation} 
that as long as the truncation is greater than a few virial radii, the choice of truncation
radius does not broadly affect the scatter on $\kapparay-\kapparec$; from now on we choose 
to truncate the halos at $5 \Rhalo$. Since $\sigma\left(\kapparay-\kapparec\right)$ does not go to
zero, it is clear that out halo based prescription does not perfectly replicate the n-body results, we refer to this
 as the intrinsic error of the prescription.
 The intrinsic error has a standard deviation of $\sigi \sim 0.018$, although
this figure is inflated by a few outliers in the tails of $\pr\left(\kapparay-\kapparec\right)$. Such outliers would 
likely be rejected by their reconstructed shear not matching the shear of the lens model 
(see \citet{wong} for the practicalities of this);
 this makes our estimate of the intrinsic error pessimistic. We do not include any additional constraints on P($\kappax$), given prior knowledge of the external shear in this work.

\begin{figure}
\includegraphics[width=\columnwidth]{truncation_scatter.eps}
\caption[magcut]{The standard deviation in $\kapparec$ minus $\kapparay$ as a function of halo truncation radius in virial radii, using the truncated NFW profile of \citet{BMO}. In our models we truncate our halos at 5$R_{200}$}
\label{fig:ScattervsTruncation}
\end{figure}

For $\kapparec$ to provide useful constraints on $\kapparay$, it is important that $\kapparay$ and $\kapparec$ are as similar as possible. Figure \ref{fig:bias} shows that indeed $\kapparec$ is a good estimator, regardless of the individual value of $\kapparec$, but that the scatter increases with $\kapparec$. The curve in $\kapparec-\kapparay$ at high $\kapparec$ shows that the reconstruction does not perfectly reproduce the physics of the most overdense lines of sight. Since most lines of sight have $\kappax<0.05$, $\kapparec$ is an acceptable estimator of $\kapparay$, although there is some scope for improvement at the highest $\kappax$.

\begin{figure}
\includegraphics[width=\columnwidth]{isitbiased.eps}
\caption[Biased?]{$\kapparec$ minus $\kapparay$ as a function of $\kapparec$, for 10000 reconstructed lines of sight. The points are put into bins of width 0.01 in $\kapparec$, except the last bin containing all reconstructions with $\kapparec>0.14$. The errorbars show the population standard deviation for each bin. Since all of the errorbars are consistant with zero, there is no evidence for $\kapparec$ being a biased estimator of $\kapparay$}
\label{fig:isitbiased}
\end{figure}

The intrinsic error between the results of the halo based prescription and the raytraced convergence is an unavoidable error, even with perfect knowledge of halo mass and position. Before investigating the errors caused by imperfect knowledge of halo mass and redshift, we investigate the errors induced by a magnitude limited reconstruction and a field of view limited reconstruction. The halos in our catalogue are given magnitudes by the semi-analytic model of \citet{SAM} and by applying magnitude cuts to our catalogue, we can investigate the scatter caused by unobserved halos. The majority of the convergence comes from halos with an $i$ magnitude between 18 and 24. Figure \ref{fig:magcut} shows that the width of $\kapparec-\kapparay$ decreases quickly between $i=18$ and $i=24$. Objects brighter than $i=18$ are either too rare or too close to the observer to make a significant contribution to the convergence. Objects fainter than $i=24$ are too small to be important unless they are extremely close to the line of sight where it is likely that neglecting stellar mass and using a spherical NFW prescription are too naive to adequately reconstruct $\kappax$ - it is likely that these ultra-faint halos (or substructures), the uncertain contribution of voids and halos deviating from our spherical truncated NFW profile are the three major sources of the scatter in $\kapparec-\kapparay$; a deeper survey will not be sufficient to decrease the size of this scatter. Since a large number of callibration lines of sight must be reconstructed to sample $\pr(\kappav)$, it may prove unrealistic to reconstruct the calibration lines of sight out to 5 arcminutes. Figure \ref{fig:radcut} shows the width of $\kapparec-\kapparay$ as a function of reconstruction radius; in almost all cases there is no significant contribution to $\kappa$ from halos that are centred more than 1 arcminute away, although large groups or clusters can sometimes still make a contribution as far out as four arcminutes. For the rest of this work we will continue to model all of the halos out to 5 arminutes with an $i$ magnitude greater than 26, although a reconstruction with fields going out to two arcminutes and down to $i=24$ would do almost as well.

\begin{figure}
\includegraphics[width=\columnwidth]{mag_scatter.eps}
\caption[magcut]{The 16, 50 and 84\% confidence intervals on $\kapparec$ minus $\kapparay$ as a function of the limiting $i$ band depth of the halo reconstruction - there is no significant improvement from including small halos with magnitudes less than $i=24$}
\label{fig:magcut}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{radius_scatter.eps}
\caption[radius cut]{The 16, 50 and 84\% confidence intervals on $\kapparec$ minus $\kapparay$ as a function of the limiting radius (in arcminutes) beyond which halos are not reconstructed. The majority of the convergence comes from halos centred within an arcminute of the line of sight}
\label{fig:radcut}
\end{figure}

The final test that we apply to the perfect knowledge reconstruction is to apply the marginalization outlined in Section \ref{subsec:voids}. Taking the reconstructed $\sumkappah$ for $1\times 10^{-5}$ lines of sight we form the joint distribution, P$(\kapparay,\sumkappah)$ and marginalize over all sightlines with a similar $\sumkappah$ to give P($\kappax$). We define the bias as the difference between the expectation value of $\kappax$ and the known true value of $\kapparay$ for each Line of sight. From $1\times 10^{4}$ reconstructed lines of sight we find the the bias to have mean of $2.0\times 10^{-5}$ and standard deviation of 0.0193. In the case where one assumes P$(\kappax)$ is given by the global P$^{\rm global}(\kappax)$, the same $1e4$ lines of sight show a mean bias of $7.7\times 10^{-4}$ and the bias has a standard deviation of 0.0346. Within the statistical error both methods are unbiased, but for any individual line of sight the bias is typically 1.8 times smaller with the reconstructed P$^{\rm reconstruction}(\kappax)$, rather than global P$^{\rm global}(\kappax)$.

@@

Real attempts to reconstruct the convergence along a line of sight will not have direct access to the masses of dark matter halos, 
and will likely not have access to their spectroscopic redshift either. Instead these properties must be inferred from astrophysical
observables. In principle the only observables are astrometric positions, photometric colours and possibly spectra. In this section we
quantify the errors induced by inferring the halo mass and redshift from the observables. Much work has already focused on using
photometric colours to infer stellar mass \citep[\eg][]{xxx} and redshifts \citep[\eg][]{BPZ}, in line with these we shall use
stellar mass and photometric redshifts (with appropriate errors), as the observables that must be converted into an estimate
of $\kappax$. We will investigate two main sources of error: inferring halo mass given observed stellar mass and the combination of stellar mass errorsand placing halos at the wrong redshift due to photometric redshift error.

We generate a stellar mass for each halo according to the stellar mass--halo mass relation of \citet{behroozi2010}. From this stellar mass we simulate 
an observed stellar mass by drawing a sample from $\pr(\log(M_{* \mathrm {obs}})|\log(M_{* \mathrm {true}}))$ which is given by the product of the galaxy stellar mass function (GSMF) of \citet{Fontana2006} and a Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where a spectroscopic redshift exists stellar masses can be estimated with typical uncertainties of 0.15 dex \citep{xxx}, however with photometric redshifts stellar mass uncertainties are typically three times as large \citep{xxx}; we use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and $\sigma_{M_*}=0.45$ otherwise. The \citet{Fontana2006} GSMF is given by a Schechter function with redshift evolving parameters
\be
\label{eq:GSMF}
{\dee N \over \dee M_*} \propto \left(10^{(M_*-M_0(z))}\right)^{(1+\alpha(z))} \exp \left(-10^{(M_*-M_0(z))}\right)
\ee 
where $M_0(z)=11.16+0.17z-0.07z^2$ and $\alpha(z)=-1.18-0.082z$. For photometric redshift errors we draw a redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take as a Gaussian of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$ is the halo's true redshift in the \MS catalogue

For each investigated source of error, we reconstruct 10000 lines of sight and each of the reconstructed $\pr(\sumkappah)$ is summarized by their mean $\sumkappah$, and their half-width containing $68\%$ of the probability, $\sigt$.
Reconstructing the lines of sight with perfect knowledge of the redshift but an uncertain stellar mass, we find that $\sigt$ grows with $\sumkappah$; this is not surprising since the low $\sumkappah$ lines of sight are relatively empty and so there are few opportunities for uncertainties in the halo masses to propogate into $\sumkappah$ uncertainties. The means and widths of these reconstructions are summarized in Figure \ref{fig:width1}, before and after applying the void correction and global prior ${\rm U}(\kappax)$. Before applying the void correction, the width of the reconstructions are roughly proportional to the mean of $\sumkappah$; they are well fit by a straight line given by 
\be
\sigt=-0.003+0.25 \langle\sumkappah\rangle.
\label{eq:Mstarfit}
\ee
After applying the void correction the expected bias is $-1.7\times 10^{-4}$ and the standard deviation for the bias is given by 0.0231; this is 1.2 times worse than the reconstruction given perfect knowledge of the halo masses and redshifts but still 1.5 times better than using the global P$^{\rm global}(\kappax)$; this is the reconstruction error that could be expected from a complete spectroscopic survey of the field. A complete spetroscopic survey is unrealistic, but Equation \ref{eq:Mstarfit2} represents a lower limit on the reconstruction error. For a randomly selected line of sight, the expectation value for the uncertainty on the reconstruction is 0.017 and the expectation value for the bias on E$(\kapparec)-\kapparay$ is 0.004.

\begin{figure}
\includegraphics[width=\columnwidth]{kappaspectro.eps}
\caption[radius cut]{Means and widths of $\kapparec$ for a complete spectroscopic survery of all halos down $to i=26$ and within 5 arcminutes of the line of sight. Results for 10000 lines of sight are shown. Red is the width and mean after summing up all the reconstructed halo contributions. Blue is the width of mean after correcting for voids. Green is the result after weighting the PDFs so that the universal kappa distribution is recovered}
\label{fig:width1}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{kappasmall.eps}
\caption[radius cut]{As figure \ref{width1}, but for a photometric reconstruction of all halos down $to i=24$ and within 2 arcminutes of the line of sight}
\label{fig:width2}
\end{figure}
